---
title: "Working Draft"
author: "Abrar Trabulsi"
date: "4/20/2020"
output: bookdown::pdf_document2
#bibliography: finalbib.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
#knitr::write_bib(c("knitr", "stringr"), "finalbib.bib", width = 60)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(texreg)
library(rstanarm)
library(stargazer)

# **** REMEMBER TO CITE EVERYTHING AND TO FOOTNOTE WHERE NEEDED!!! REREAD THE RUBRIC TO KNOW FOR SURE WHERE THIS NEEDS TO HAPPEN
```
$$\textbf{Abstract}$$
  In his paper 'The Desire for Social Status and Economic Conservatism among Affluent Americans', Thal (2020) shows that affluent American's desire for social status drives conservative attitudes amongst them, and the will to advance economically conservative politics. Overall, I was successful in my replication efforts in this paper. Moreover, I extend Thal's results by running logistic regression and appropriate analysis, such a distribution of the posterior and predictive probabilities on his primary data. I find that Thal was indeed correct. Social status does in fact drive economically conservative attitudes amongst affluent Americans, and especially men.
$$\textbf{Introduction}$$
  Thal begins by establishing that affluent Americans have considerable power over policy outcomes, and oftentimes use their influence to favor themselves, by advancing conservative economic policies. His study addresses a lack in relevant literature which discusses affluent Americans successfully pushing economic policy to the right, by showing evidence that the reasons behind this are psycho-social. Thal does this with a dual layered methodology. First he establishes through a survey that Americans' desire for social status strongly predicts their economic conservatism. Next, through an experiment conducted on social media, he shows that competition for social status drives economic conservatism. Thal's study builds on literature by Veblen (1899) which first documented such status seeking behavior. It however, contradicts studies such as that of Sears et.al (1980) which posits that affluent Americans support and advance conservative economic policies in order to improve their direct and material wealth. Thal finds no evidence during his study which supports this, and could hint either at the emphasized importance of psycho-social factors such as competition and the desirability of status, or, the increaisng importance of social media which makes displays of status and competition all the more important in our society. 
  The following paper is first and formost a replication paper with an extension of Thal's results. All replication was conducted in R with the original code found in the Harvard Data verse HERE as well as my repo HERE. All code related to this paper and the extension will also be found at this repo.
  The primary and independent findings of this paper revolved around my extension and Thal's results. I decided to run logistic regression on Thal's main data, with the same regression variables as Thal's OLS model. I decided to run logistic regression due to the practicality of interpreting information of this nature probabilistically, in my opinion it's much more meaningful and comprehessible than in point values. I also provide the posterior and posterior predictive probability distributions of this regression, with brief but substantive analysis in the caption. Through my extension, I found that Thal was indeed correct. Social status does drive economically conservative attitudes amongst affluent Americans, and especially affluent men. Thal's results being corroborated by an entirely different model greatly validates them, and only highlights the merit of his hypothesis, the rigor of his study, and data, and the necessity of further exploring this subject as well as examining the implications of these findings. 
  
$$\textbf{Review of Literature}$$
  Studying the relationship between affluent Americans and money is nothing new. Thorstein Veblen (1899) was the first to argue that affluent Americans pursue money for the sake of social status in his paper Theory of the Leisure Class. This pursuit is driven by two distinct phenomenon according to literature: Status motivation and concrete motivation (Lea and Webley 2006; Srivastava, Locke, and Bartol 2001; Zelizer 1994). Status motivation is defined as pursuing money for the sake of esteem, which Velbin recognized as an essential part of upper-class culture. Concrete  motivation on the other hand, is defined as the pursuit of money for material needs. Thal emphasizes that these phenomenon con coexist, but are empirically distinct which he provides evidence for in a following section. 
  How does this behavior influence the population as a whole, however? Certainly, affluent Americans have a far greater ability to influence legislation in any which way they desire from the pure fact that they have more capital than a middle class or low income American. From campaign contributions to lobbying, affluent Americans have far greater options and levrage to advocate for the legislation they desire. Following the logic established above, it wouldn't be ridiculous to claim that affluent Americans are then more likely to advocate on behalf of conservative economic policies, because after all, those are the policies most likely to secure them capital, and therefore status and material security. More accurate than implicating all affluent Americans however, is to specifiy this claim to affluent men in particular. Men after all to comprise the majority of legislative offices, function as the heads of most lobbiest and interest groups, as well as donate more money to campaigns than their female counterparts (Weber et al. 2015;Wolfers 2015). It would be reasonable to say they share most of the responsibility for the rise of conservative economic policy in the United States which benefits so few, and harms so many. 
  Thal's research connects and builds on existing literature in the field by showing an empirical link between this pursuit of affluence for the sake of social status drives economic conservative attitudes and advocacy. More cleverly so, half of his study is conducted through social media. Social media is an ever present medium for communication and information in the dialy lives of Americans, and apps such as Instagram do allow those who have it, to flaunt their affluence and status quite openly. The ever presence of social media means an ever presence of these displays of affluence and status, and it would not be a stretch to say that this in of itself could further drive these behaviors amongst the affluent, or at the very least, that social media is a medium that deserves sholarly exploration in its own right. Thal theorizes that the desire for social status through affluence drives conservative economic advocacy. He tests his thoery with both an observational and experimental approach. With a rigorous survey scale and a representative sample of affluent Americans, he shows that there in fact is a link between affluent Americans desires for status and support for economically conservative policy. He also finds that this relationship is especially stong in affluent men. With a seconf sample of affluent Americans, he observes their behavior when encountering posts on Facebook expressing sentiments of self worth and status following career achievments, as well as posts showing the lack thereof. After observing differences in behavior and attitudes towards the posts in the form of likes, comments etc, patticipants were asked to answer some reinforcing questions. After this part of the study, Thal concludes the same as he did after the observational pieve of his study: that desires for affluence to achieve status drive conservative economic policy support in affluent Americans, and especially men. 
  Ultimately, this study filled a gap in existing literature and linked two separate theories regarding affluent Americans and thier drive for more affluence, as well as their political behavior. Furthermore, Thal established that the drive for affluence and subsequent political behavior is primarily driven by desires for status, not for material well being. This has great implications because the same money spent to attain more status with the purchase of a new property, coture, or other status symbols could have easily been used to help provide more and better public goods. Lastly, Thal also provides evidence that social media does in fact impact political behavior which can increase the levels of inequality in America, which are already staggering. Thal claims that he is the first to do so. 
$$\textbf{Replication}$$
  I was able to successfully able to replicate all relevant tables and figure from this study.


$$\textbf{Extension}$$
  The extension I conducted for this paper comprised of running a logistic regression on the main model, the results of which are found in Table 4, as well as plotting the posterior, and posterior predictive probabilities. 
  Table 4 contains the results of the main model of 'Test 2' of this paper (the experiment portion), an OLS regression displaying the relationship between economic conservatism as well as the interaction variable of male, each status treatment, as well as the 'concrete' condition. Table 4 can be found below.
  
```{r Table 4 Ex}
data2 <- read_rds("../finalproject1006/readata/data2.rds")

aff.lm <- lm(econcon ~ condition, data = data2)
aff.pval <- pt(coef(summary(aff.lm))[, 3], summary(aff.lm)$df[2], lower=FALSE)

aff.lmg <- lm(econcon ~ condition*male, data = data2)
aff.pvalg <- pt(coef(summary(aff.lmg))[, 3], summary(aff.lmg)$df[2], lower=FALSE)

texreg(list(aff.lm, aff.lmg),
			 override.pvalues = list(c(aff.pval),
			                         c(aff.pvalg)),
       caption.above = TRUE, stars = c(0.001, 0.01, 0.05),
			 custom.model.names = c("Main Model", 
			                        "Gender Interaction Model"),
       custom.coef.names = c("Intercept",
        										"Concrete",
       											"Conspicuous Consumption",
       											"Self-Esteem",
       											"Social Approval",
        										"Affluent Male",
        										"Concrete X Affluent Male",
       											"Conspicuous Consumption X Affluent Male",
       											"Self-Esteem X Affluent Male",
       											"Social Approval X Affluent Male"))
```

  I decided on logistic regression because I found it interesting to be able to interpret these results probablistically, and believe that in the context of this study, probabilities are more widely understandable than point values. 
  The logistic regression table is below: 
```{r Extension Regression, cache = TRUE}
mydata <- data2 %>%
  mutate(econcon = ifelse(econcon > .5, 1, 0))

fit1 <- stan_glm(econcon ~ condition*male, data = mydata, refresh = 0, family = "binomial")

print(fit1, digits = 4)
```


  To conduct this regression I had to make an adjustment to Thal's data. The 'econcon' or economic conservative variable, was originally calculated as the mean of 4 other variables which correlated to questsions asked to the participants after being exposed to a status treatment. The responses to those 4 questions (and variable values) were on a 0-1 scale, with < .5 being not economically conservative, and > .5 being economically conservative. In order to analyse these data logistically, it was necessary that I recode the econcon variable as either 0 or 1, with the original scales for the individual variables applying (> .5 is economically conservative and vice versa).
  I will present results analogus to those which Thal discussed in his paper. Thal idetifies 'Social Approval' and "Conspicuous Consumption' as the status treatments with significant effect of affluent American's economic conservatism. I find that the expected median probability difference between men and women testing as economically conservative after being exposed to the 'Conspicuous Consumption' and 'Social Approval' status treatment  is .5 and .51 respectively; men are more likely to test economically conservative after these treatments. The statistically insignificant results for the 'Self Esteem' status treatment defied Thal's original hypothesis, and in my logistic analysis I find that the expected median probability difference between men and women testing as economically conservative after being exposed to the 'Self Esteem' treatment is .44, less than the those found for the other two status treatments. 
  Lastly, the interaction variables for each statistically significant status treatment (according to Thal) can be interpreted as the median difference in the probability of a male testing economically conservative when having been treated with a particular status treatment, versus the others. According this this, a male who recieved the 'Social Approval' status treatment has an expected median probability of .64 to test as economically conservative compared to the other status treatments. This validates Thal's hypothesis that social status, which is akin to social approval, motivates affluent Americans, especially men, to support conservative economic policies. 
  The last installment of this section shows the poterior probability distribution, as well as the posterior predictive distribution for these data, specifically analyzing the probabilities of testing as economically conservative amongst men who were treated with either the 'Social Approval' or 'Conspicuous Consumption' status treatments. More quantitative interrpretations can be found in the captions.
```{r calculating propabilities, include = FALSE}
plogis(.0027)

# median likelihood of someone exposed to conspicuous consumption tests as economically conservative

plogis(.0727)

# median likelihood of someone exposed to social approval tests as economically conservative

plogis(-.2116)

# median likelihood of someone exposed to self-esteem post condition tests as economically conservative

plogis(.5803)

# likelihood that a male with social approval post treatment applied tests as economically conservative

plogis(.5483)

# lielihood that a male with conspicuous consumption post treatment applied tests as economically conservative
```

```{r Extension Posterior Dist}
new = data.frame(male = 1, condition = c("Conspicuous Consumption","Social Approval"))

m1 <- posterior_linpred(fit1, transform = TRUE, newdata = new)

hist(m1)

# the distribution of posterior probabilities of testing as economically conservative if male and either conspicuous consumption or social approval post treatment is applied
```
```{r Extension Posterior Predictions}
m2 <- posterior_predict(fit1, transform = TRUE, newdata = new)

hist(m2)

# predictions of testing economically conservative if you're a male who was treated with either Conspicuous Consumption or social approval given the posterior probabilities
```
  
$$\textbf{Conclusion}$$



$$\textbf{Appendix}$$
Below you'll find all replicable Figures and Tables in the order they appear in the paper:
*Please look at the RMD file found HERE for the code to replicate these tables and figures

# Figure 1

```{r Figure 1, include = FALSE}
df <- read_rds("../finalproject1006/readata/df.rds")

ggplot(df, aes(x=Concrete, y=Status)) +
  theme_bw() +
  labs(x = "PC1: Concrete", y = "PC2: Status") +
  scale_x_continuous(limits = c(-1,1), breaks = seq(-1, 1, .25)) +
  scale_y_continuous(limits = c(-1,1), breaks = seq(-1, 1, .25)) +
  coord_cartesian(xlim=c(-.95, .95), ylim=c(-.95, .95)) +
  geom_segment(aes(x=-1,xend=1,y=0,yend=0), colour="dark grey", linetype="solid") +
  geom_segment(aes(x=0,xend=0,y=-1,yend=1), colour="dark grey", linetype="solid") +
  geom_text(data=df, aes(x=Concrete, y=Status, 
                         label = Motive, fontface = "bold"),
            size = 6) +
  theme(text = element_text(size=25),
        plot.margin=unit(c(1,1,1,1.2),"cm"))
```

# Figure 2

```{r Figure 2, include = FALSE}
data <- read_rds("../finalproject1006/readata/data.rds")

ggplot() +
  geom_point(data=data, aes(x=concrete, y=status), size = 3, color = 'grey') +
  theme_bw() +
  labs(x = "Concrete Component Score", y = "Status Component Score") +
  scale_x_continuous(breaks = seq(-3, 3, .5),
                     sec.axis = dup_axis(name = "", 
                                         breaks = seq(-3, 3, length.out = 9),
                                         labels = sprintf("%.2f", seq(-1, 1, .25)))) +
  scale_y_continuous(breaks = seq(-3, 3, .5),
                     sec.axis = dup_axis(name = "", 
                                         breaks = seq(-3, 3, length.out = 9),
                                         labels = sprintf("%.2f", seq(-1, 1, .25)))) +
  coord_cartesian(xlim=c(-2.95, 2.95), ylim=c(-2.95, 2.95)) +
  geom_segment(aes(x=-3,xend=3,y=0,yend=0), colour="grey", linetype="dashed") +
  geom_segment(aes(x=0,xend=0,y=-3,yend=3), colour="grey", linetype="dashed") +
       theme(text = element_text(size=30),
        		 axis.text.x=element_text(size=20),
        		 axis.text.y=element_text(size=20)) +
  geom_text(data=df, aes(x=Concrete.p, y=Status.p, 
                         label = Motive, fontface = "bold"),
            						size = 5.5) +
   geom_segment(data=df, aes(xend = Concrete.a,yend=Status.a),
              x=0, y=0, colour="black",
              arrow=arrow(angle=25, length=unit(0.5, "cm")),
              linetype="dashed",
   						size=1) 
```

# Table 2

```{r Table 2, include = FALSE}
dataf1 <- read_rds("../finalproject1006/readata/dataf1.rds")

econcon.aff.lm <- lm(econcon ~ 
					      status_2sd +
					      concrete_2sd +
					      COLI_2sd +
                male + 
                asian +
					      hispanic +
					      nhblack +
					      other +
                age3044 + 
                age4554 + 
                age55up + 
                college + 
                grad, 
					      weights = weight, 
                data = dataf1)

econcon.aff.lmg <- lm(econcon ~
					      status_2sd*male +
					      concrete_2sd*male +
					      COLI_2sd +
                asian +
					      hispanic +
					      nhblack +
					      other +
                age3044 + 
                age4554 + 
                age55up + 
                college + 
                grad, 
					      weights = weight, 
                data = dataf1)

texreg(list(econcon.aff.lm, econcon.aff.lmg),
       single.row = FALSE,
       caption.above = TRUE, stars = c(0.001, 0.01, 0.05), 
       custom.model.names = c("Main Model", "Gender Interaction Model"),
       custom.coef.names = c("Intercept",
        										"Status",
        										"Concrete",
        										"Regional Cost of Living",
        										"Male",
        										"Asian",
        										"Latino",
        										"Black",
        										"Other",
        										"Age 30-44",
        										"Age 45-54",
        										"Age 55 and up",
        										"College degree",
        										"Graduate degree",
														"Status X Male",
														"Concrete X Male"))

```

# Figure 3

```{r Figure 3, include = FALSE}
df3 <- read_rds("../finalproject1006/readata/df3.rds")

ggplot(df3, aes(x = level, y = pred, ymin = lwr, ymax = upr, 
							 shape = type, group = type, colour = type, linetype = type)) + 
	geom_errorbar(width=.1, size=1, position = "dodge", linetype = "solid") +
  geom_point(size=5, position= "dodge", aes(colour = type)) +
  geom_line(aes(group=type), position="dodge") +
	theme_bw() +
  scale_colour_manual(values=c("grey35", "grey65"))  +
  scale_linetype_manual(values=c("solid", "dashed"))  +
  ylab("Economic Conservatism") +
	xlab("Level") +
  scale_y_continuous(limits=c(.45, .75),
                     breaks = seq(0,1,.05)) +
  theme(text = element_text(size=25),
        axis.text.x=element_text(size=25)) +
	guides(shape = guide_legend(reverse=F, title="Motive"),
				 colour = guide_legend(reverse=F, title="Motive"),
				 linetype = guide_legend(reverse=F, title="Motive"))
```

# Figure 4

```{r Figure 4, include = FALSE}
df4 <- read_rds("../finalproject1006/readata/df4.rds")

ggplot(df4, aes(x = level, y = pred, ymin = lwr, ymax = upr, 
							 shape = type, group = type, colour = type, linetype = type)) + 
  facet_grid(.~sex) +
	geom_errorbar(width=.1, size=1, position="dodge", linetype="solid") +
  geom_point(size=5, position="dodge", aes(colour = type)) +
  geom_line(aes(group=type), position="dodge") +
	theme_bw() +
  scale_colour_manual(values=c("grey35", "grey65"))  +
  scale_linetype_manual(values=c("solid", "dashed"))  +
  ylab("Economic Conservatism") +
	xlab("Level") +
  scale_y_continuous(limits=c(.45, .75),
                     breaks = seq(0,1,.05)) +
  theme(text = element_text(size=25),
        axis.text.x=element_text(size=25)) +
	guides(shape = guide_legend(reverse=F, title="Motive"),
				 colour = guide_legend(reverse=F, title="Motive"),
				 linetype = guide_legend(reverse=F, title="Motive"))

```

# Table 4

```{r Table 4, include = FALSE}

aff.lm <- lm(econcon ~ condition, data = data2)
aff.pval <- pt(coef(summary(aff.lm))[, 3], summary(aff.lm)$df[2], lower=FALSE)

aff.lmg <- lm(econcon ~ condition*male, data = data2)
aff.pvalg <- pt(coef(summary(aff.lmg))[, 3], summary(aff.lmg)$df[2], lower=FALSE)

texreg(list(aff.lm, aff.lmg),
			 override.pvalues = list(c(aff.pval),
			                         c(aff.pvalg)),
       caption.above = TRUE, stars = c(0.001, 0.01, 0.05),
			 custom.model.names = c("Main Model", 
			                        "Gender Interaction Model"),
       custom.coef.names = c("Intercept",
        										"Concrete",
       											"Conspicuous Consumption",
       											"Self-Esteem",
       											"Social Approval",
        										"Affluent Male",
        										"Concrete X Affluent Male",
       											"Conspicuous Consumption X Affluent Male",
       											"Self-Esteem X Affluent Male",
       											"Social Approval X Affluent Male"))
```

# Figure 6

```{r Figure 6, include = FALSE}

plot <- read_rds("../finalproject1006/readata/plot.rds")

ggplot(plot, aes(x=condition, y=econcon)) + 
    geom_errorbar(aes(ymin=econcon-ci, ymax=econcon+ci), 
                  width=.15, position = position_dodge(width = 0.5), size = .75) +
    geom_point(position = position_dodge(width = 0.5), size=6,
               shape=18) + 
	  geom_text(aes(label=label),hjust=-.25, vjust=.35, size=8) +
    ylab("Economic Conservatism") +
	  xlab("") +
    theme_bw() +
       theme(text = element_text(size=20),
          legend.position="none",
          axis.text.x = element_text(size = 25),
          axis.text.y = element_text(size = 25),
    			axis.title.y = element_text(size = 25)) +
    scale_y_continuous(limits=c(.55, .75),
                      breaks = seq(.55,.75,.05),
                      labels = c(".55",".60",".65",".70",".75")) +
    scale_x_discrete(labels = c("Placebo",
                                "Conspicuous\nConsumption",
                                "Self-\nEsteem",
                                "Social\nApproval",
    														"Concrete"))
```

# Figure 7

```{r Figure 7, include = FALSE}
plotm <- read_rds("../finalproject1006/readata/plotm.rds")

ggplot(plot, aes(x=condition, y=econcon)) + 
    geom_errorbar(aes(ymin=econcon-ci, ymax=econcon+ci), 
                  width=.15, position = position_dodge(width = 0.5), size = .75) +
    geom_point(position = position_dodge(width = 0.5), size=5,
               shape=16) + 
	  geom_text(aes(label=label),hjust=-.25, vjust=.35, size=8) +
    ylab("Economic Conservatism") +
	  xlab("") +
    theme_bw() +
       theme(text = element_text(size=20),
          legend.position="none",
          axis.text.x = element_text(size = 25),
          axis.text.y = element_text(size = 25),
    			axis.title.y = element_text(size = 25)) +
    scale_y_continuous(limits=c(.55, .8),
                      breaks = seq(.5,.8,.05),
                      labels = c(".50", ".55",".60",".65",".70",".75", ".80")) +
    scale_x_discrete(labels = c("Placebo",
                               "Conspicuous\nConsumption",
                               "Self-\nEsteem",
                               "Social\nApproval",
    													 "Concrete"))

plotw <- read_rds("../finalproject1006/readata/plotw.rds")

ggplot(plot, aes(x=condition, y=econcon)) + 
    geom_errorbar(aes(ymin=econcon-ci, ymax=econcon+ci), 
                  width=.15, position = position_dodge(width = 0.5), size = .75) +
    geom_point(position = position_dodge(width = 0.5), size=5,
               shape=17) + 
	  geom_text(aes(label=label),hjust=-.25, vjust=.35, size=8) +
    ylab("Economic Conservatism") +
	  xlab("") +
    theme_bw() +
       theme(text = element_text(size=20),
          legend.position="none",
          axis.text.x = element_text(size = 25),
          axis.text.y = element_text(size = 25),
    			axis.title.y = element_text(size = 25)) +
    scale_y_continuous(limits=c(.55, .8),
                      breaks = seq(.55,.8,.05),
                      labels = c(".55",".60",".65",".70",".75", ".80")) +
    scale_x_discrete(labels = c("Placebo",
                               "Conspicuous\nConsumption",
                               "Self-\nEsteem",
                               "Social\nApproval",
    													 "Concrete"))
```

$$\textbf{Bibliography}$$

